General constraints on magic distillation protocols from statistical mechanics

### 2021-06-03 magic_paper.tex
fixed majorization to relative majorization proof

\subsection{Schur-Convexity and quasi-distributions}
We now make use of known results from majorization theory. We first define \emph{Schur-convex functions} on a subset $\D \subseteq \mathbb{R}^N$.

\begin{definition} A continuous, real-valued function $f$ defined on $\D \subseteq \mathbb{R}^N$ is called Schur-convex on the domain $\D$ if $\p \prec \q$ on $\D$ implies that $f(\p) \le f(\q)$. A continuous function $f:\D \rightarrow \mathbb{R}$ is called Schur-concave if $(-f)$ is Schur-convex.
\end{definition}
We now can consider the set of all quasi-distributions, which form a hyper-plane and are given by
\begin{equation}
\D = \{ \w \in \mathbb{R}^N: \sum_i w_i = 1\}.
\end{equation}
For functions that are symmetric, namely $f(\p) = f(\Pi \p)$ for any component permutation $\Pi$, it suffices to restrict to the following region,

\begin{equation}
\D^\downarrow = \{ \w \in \D : w_1 \ge w_2 \ge \cdots \ge w_N\}.
\end{equation}

We now have the following theorem~\cite{cit:marshall} that provides a characterization of symmetric, Schur-convex functions on quasi-distributions.
\begin{theorem} Let $f$ be a real-valued continuous function defined on $\D^\downarrow$ that is continuously differentiable on the interior of $\D^\downarrow$. Then $f$ is Schur-convex on $\D^\downarrow$ if and only if $\partial_k f(\w)$ is decreasing in $k$ for all $\w \in \D^\downarrow$, namely
\begin{equation}
\partial_1 f(\w) \ge \partial_2 f(\w) \ge \cdots \ge \partial_N f(\w),
\end{equation}
for all $\w$ in the interior of $\D^\downarrow$.
\end{theorem}

\begin{lemma}[Additivity of $H_{2m}$]\label{H_add}
	For any states $\rho, \rho'$ and non-negative $m$, we have
	\begin{equation}
		H_{2m}(\rho \otimes \rho') = H_{2m}(\rho) + H_{2m}(\rho')
	\end{equation}
\end{lemma}
\begin{proof}
	This can be seen directly by the multiplicative property of the Wigner distribution,
	\begin{align*}
		H_{2m}&(\rho \otimes \rho') = \frac{1}{1-2m}\log \sum_{\bmx,\bmx'} \left[W_{\rho}(\bmx)W_{\rho}(\bmx')\right]^{2m} \\
		&= \frac{1}{1-2m}\left( \log \sum_{\bmx}W_{\rho}(\bmx)^{2m} + \log \sum_{\bmx'}W_{\rho}(\bmx')^{2m} \right) \\
		&= H_{2m}(\rho) + H_{2m}(\rho').
	\end{align*}
\end{proof}
\cref{H_add} implies that for any $n$,
\begin{equation}
	H_{2m}(\rho_n) := H_{2m}(\rho^{\otimes n}) = n H_{2m}(\rho).
\end{equation}

@@

For an $\epsilon$-noisy Strange state, we have
\begin{equation}
	H_{\alpha}\left(\rho_{\rm{S}}(\epsilon)\right) = \frac{1}{1-\alpha}\log_2 \left[ 8\left( \frac{1}{6} - \frac{1}{18}\epsilon \right)^\alpha + \left(-\frac{1}{3} + \frac{4}{9}\epsilon \right)^\alpha \right]
\end{equation}

We also have for any quantum state $\rho$ that
\begin{equation}
	H_{0}\left(\rho\right) = 2\log_2 3 = H_\alpha \left( \id/3 \right),
\end{equation}
for any $\alpha$, so the bound is not well-defined for $H_0$. \nick{this should be true for general relative majorization.}
\subsection{Dropping pre/post-processing of the magic states}
\label{app:cliff_processing}

>>>>

\subsection{Entropic bounds on quasi-distributions}

We first define the classical $\alpha$-Renyi entropy in terms of the Wigner quasi-distribution of any state $\rho$ as
\begin{equation}
	H_\alpha(\rho) := \frac{1}{1-\alpha} \log \sum_{\bmx} W_{\rho}(\bmx)^\alpha,
\end{equation}
where $\alpha$ can take values of the form $\alpha_1 / \alpha_2$, with even $\alpha_1 \geq 0$ and odd $\alpha_2 > 0$.

We then prove the following result connecting simple majorization to relative majorization.
\nick{move following lemma to section on relative majorization equivalent conditions}
\begin{lemma}
Let $\w\in \mathbb{R}^N, \w'\in \mathbb{R}^{N'}$ be quasi-distributions and $\r\in \mathbb{R}^N, \r'\in \mathbb{R}^{N'}$ probability distributions with positive rational entries given by $r_i = a_i/K$ and $r_i' = a_i'/K$ for positive integers $a_i, a_i'$ and $K = \sum_{i=1}^N a_i = \sum_{i=1}^{N'} a_i'$. 
Then,
\begin{equation}
(\w, \r) \succ (\w', \r') \mbox{ if and only if } \Gamma_{\bma} (\w) \succ \Gamma_{\bma'} (\w'),
\end{equation}
where the embedding map $\Gamma_{\bma} :\mathbb{R}^N \rightarrow \mathbb{R}^K$ is given by
\begin{equation}
	\Gamma_{\bma}(\x) := \bigoplus_{i=1}^N x_i \eta_{b_i},
\end{equation}
with $\eta_{a_i} = (1/a_i, 1/a_i, \dots, 1/a_i)$ the uniform distribution on $a_i$ elements.
\end{lemma}
\begin{proof}
	We first note that the map $\Gamma_{\bma}$ is stochastic and has a well-defined left-inverse $\Gamma^{-1}_{\bma'}: \mathbb{R}^K \rightarrow \mathbb{R}^N$ given by
\begin{equation}
	\Gamma^{-1}_{\bma}(\w) = \left( \sum_{i=1}^{a_1} w_i,  \sum_{i = a_1 +1}^{a_1 + a_2} \hspace{-5pt} w_i, \dots , \hspace{-15pt} \sum_{i=a_1 + \dots + a_{N-1}+1}^{K} \hspace{-10pt} w_i \right),
\end{equation}
that obeys $(\Gamma^{-1}_{\bma} \circ \Gamma_{\bma}) (\x) = \x$ for all $\x \in \mathbb{R}^N$.

The claim is equivalent to the statement that there exists a \emph{bistochastic map} $B$ sending $\Gamma_{\bma}(\w)$ to $\Gamma_{\bma'}(\w')$ if and only if there exists a stochastic map $A$ sending $\w$ to $\w'$ and $\r$ to $\r'$.

Suppose there is a stochastic map $A$ such that $A\w=\w'$ and $A\r = \r'$. 
We define $B:\mathbb{R}^K \mapsto \mathbb{R}^K$ by $B := \Gamma_{\bma'} \circ A \circ \Gamma_{\bma}^{-1}$ so that it is stochastic as a composition of stochastic maps and it preserves the uniform distribution in $\mathbb{R}^A$, since
\begin{align}
	&B(1/K, 1/K, \dots, 1/K) = \big( \Gamma_{\bma'} \circ A \circ \Gamma_{\bma}^{-1} \big) \big( \Gamma_{\bma'}(\r) \big) = \nonumber\\
	&\Gamma_{\bma'} (\r') = (1/K, 1/K, \dots, 1/K),
\end{align}
therefore $B$ is bistochastic.
Finally, $B$ maps the embedded distributions as follows,
\begin{equation}
	B \Gamma_{\bma}(\w) = \big( \Gamma_{\bma'} \circ A \circ \Gamma_{\bma}^{-1} \big) \big( \Gamma_{\bma}(\w) \big) = \Gamma_{\bma'} (\w').
\end{equation}

Conversely, suppose a bistochastic map $B$ exists sending $\Gamma_{\bma}(\w)$ to $\Gamma_{\bma'}(\w')$. Again, define $A: \mathbb{R}^N \mapsto \mathbb{R}^{N'}$ by $A := \Gamma_{\bma'}^{-1} \circ B \circ \Gamma_{\bma}$ so that it is stochastic as a composition of stochastic maps and $A \w = \big( \Gamma_{\bma'}^{-1} \circ B \circ \Gamma_{\bma} \big) \w = \w'$, as well as $A \r = \r'$.
\end{proof}
We also note that the vectors $\r, \r'$ with rational components is dense in the set of all positive probability distributions, and so we appeal to the fact that experiments have a finite resolution, and ignore technical details to do with the completion to arbitrary real components.

We now make use of known results from majorization theory. We first define \emph{Schur-convex and Schur-convave functions} on a subset $\D \subseteq \mathbb{R}^N$.

\begin{definition} A continuous, real-valued function $f$ defined on $\D \subseteq \mathbb{R}^N$ is called Schur-convex on the domain $\D$ if $\p \prec \q$ on $\D$ implies that $f(\p) \le f(\q)$. A continuous function $f:\D \rightarrow \mathbb{R}$ is called Schur-concave if $(-f)$ is Schur-convex.
\end{definition}
We can consider the set of all quasi-distributions, which form a hyper-plane and are given by
\begin{equation}
\D = \{ \w \in \mathbb{R}^N: \sum_{i=1}^N w_i = 1\}.
\end{equation}
For functions that are symmetric, namely $f(\p) = f(\Pi \p)$ for any component permutation $\Pi$, it suffices to restrict to the following region,
\begin{equation}
\D^\downarrow = \{ \w \in \D : w_1 \ge w_2 \ge \cdots \ge w_N\}.
\end{equation}

We now have the following theorem~\cite{cit:marshall} that provides a characterization of symmetric, Schur-convex and Schur-concave functions on quasi-distributions.
\begin{lemma} 
Let $f$ be a real-valued continuous function defined on $\D^\downarrow$ that is continuously differentiable on the interior of $\D^\downarrow$. Then $f$ is Schur-convex on $\D^\downarrow$ if and only if $\partial_k f(\w)$ is decreasing in $k$ for all $\w \in \D^\downarrow$, namely
\begin{equation}
\partial_1 f(\w) \ge \partial_2 f(\w) \ge \cdots \ge \partial_N f(\w),
\end{equation}
for all $\w$ in the interior of $\D^\downarrow$.
\end{lemma}

We can use the above to prove that $H_\alpha$ is Schur-concave for

\begin{lemma}[Additivity of $H_{\alpha}$]\label{H_add}
	For any $\alpha > 0$ such that $H_{\alpha}$ is well-defined and any states $\rho, \rho'$, we have
	\begin{equation}
		H_{\alpha}(\rho \otimes \rho') = H_{\alpha}(\rho) + H_{\alpha}(\rho')
	\end{equation}
\end{lemma}
\begin{proof}
	This can be seen directly by the multiplicative property of the Wigner distribution,
	\begin{align*}
		H_{\alpha}&(\rho \otimes \rho') = \frac{1}{1-\alpha}\log \sum_{\bmx,\bmx'} \left[W_{\rho}(\bmx)W_{\rho}(\bmx')\right]^{\alpha} \\
		&= \frac{1}{1-\alpha}\left( \log \sum_{\bmx}W_{\rho}(\bmx)^{\alpha} + \log \sum_{\bmx'}W_{\rho}(\bmx')^{\alpha} \right) \\
		&= H_{\alpha}(\rho) + H_{\alpha}(\rho').
	\end{align*}
\end{proof}
\cref{H_add} implies that for any $n$,
\begin{equation}
	H_{2m}(\rho_n) := H_{2m}(\rho^{\otimes n}) = n H_{2m}(\rho).
\end{equation}

@@

For an $\epsilon$-noisy Strange state, we have
\begin{equation}
	H_{\alpha}\left(\rho_{\rm{S}}(\epsilon)\right) = \frac{1}{1-\alpha}\log_2 \left[ 8\left( \frac{1}{6} - \frac{1}{18}\epsilon \right)^\alpha + \left(-\frac{1}{3} + \frac{4}{9}\epsilon \right)^\alpha \right]
\end{equation}
\ddd{[TO DO: Set $\epsilon = 0$ and determine region of $\alpha$ such that $H <0$.]}

We also have for any quantum state $\rho$ that
\begin{equation}
	H_{0}\left(\rho\right) = 2\log_2 3 = H_\alpha \left( \id/3 \right),
\end{equation}
for any $\alpha$, so the bound is not well-defined for $H_0$.

Therefore if we have a Schur-Concave function $f(\x)$ for majorization of quasi-distributions then we can define a monotonic function for relative majorization via composition with $\Gamma_{\mathbf{a}}$ and its inverse.
This should provide relative majorization monotones from which we can obtain distillation bounds, such as
\begin{equation}
R \le \frac{F_\alpha(\tau) - F_\alpha (\rho)}{F_\alpha(\tau') - F_\alpha (\rho')}.
\end{equation}
\ddd{[To be checked...transition from entropies to free energies less obvious from this route...needs reflection]}

This is probably all we need to show:
\begin{equation}
H_\alpha(\Gamma_{\mathbf{a}}(\p)) = D_\alpha(\p ||\r),
\end{equation}
where we define the Renyi relative entropy
\begin{equation}
D_\alpha (\p || \r) := \frac{1}{\alpha -1} \log \sum_i p_i^\alpha q_i^{1-\alpha}.
\end{equation}
This probably follows from the details of how the $\Gamma_{\mathbf{a}}$ relates to $\r$. Needs expanding carefully.
\subsection{Dropping pre/post-processing of the magic states}
\label{app:cliff_processing}
